Refinement for open automata

### 2021-05-06 2021-Refinement/report.tex

\begin{defi}[Non-locking composition, intuitive definition]
The composition resulting in \(A_p\subst{A_c}{k} = \OA{S}{s_0}{V}{\sigma_0}{J}{T}\) with \(A_c \defobject \OA{S_c}{s_{0c}}{V_c}{\sigma_{0c}}{J_c}{T_c}\), \(A_p \defobject \OA{S_p}{s_{0p}}{V_p}{\sigma_{0p}}{J_p}{T_p}\) and \(k \in J_p\), is a non-locking composition if there is a rechability predicate such that from reachable configurations, either a transition is possible in \(A_p\subst{A_c}{k}\) or there were no possible transition in \(A_p\) to begin with
\begin{multline*}
	\forall s \in S, \forall \sigma: V \to \values, \mpar{\sigma \vdash \reach{A_p\subst{A_c}{k}}\mpar{s}} \implies \mpar{\exists t \in \fOT{s}, \exists \nu: \fvars{t} \to \values, \sigma \uplus \nu \vdash \fguard{t}} \\
	\vee \forall t_p \in \fOT{\pi_{S_p}\mpar{s}}, \forall \nu: \fvars{t_p} \to \values, \sigma \uplus \nu \nvdash \fguard{t_p}
\end{multline*}
\end{defi}

To illustrate the introduction of a deadlock in a composition and the definition of non-locking composition, let's look at an example.
\begin{exi}[Deadlocks introduced by composition]
% TODO: give a context, a circuit cannot block physical time from progressing
\begin{figure}
\input{randomtick_clock.tex}
\vrule
\input{register_anytick.tex}
\caption{On the left: A clock which imposes a tick; On the right: A modified version of the counter at figure \ref{fig:tlh}}
\label{fig:anytick}
\end{figure}
% TODO: explain
\begin{figure}
\input{Composition_deadlock.tex}
\vrule
\input{Composition_nolock.tex}
\caption{On the left: Deadlocks introduced by composition; On the right: No deadlock introduced by composition}
\label{fig:deadlock}
\end{figure}
% TODO: explain
\end{exi}

When expanding the aliases \(S, i, V\) in the definition, and remarking that the last existential quantifier can be unfolded into a transition in \(A\) and one in \(A_p\) with the added condition that they match, the definition become mildly similar with the deadlock reduction one.
It is not a coincidence because their goal is to caracterise the same kind of compatibility between automata.
Actually it is possible to simplify both definition and make them even more similar.

\begin{defi}[Deadlock reduction, working definition] % TODO HERE: Maybe there is a union vars OT(s_2) missing in the domain of sigma
A simulation \(\leq\) is deadlock reducing if in all the \(R: S_1 \times S_2 \to \rformulae[V_1 \uplus V_2]\) such that \(\wrel{A_1}{A_2}{R}\), there is one that also satisfies
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \forall \sigma: V_1 \uplus V_2 \uplus \bigcup_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \to \values, \\
	\sigma \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}
\end{multline*}
\end{defi}
\begin{lem}
The intuitive and working definition of deadlock reduction are equivalent.
\end{lem}

>>>>

\begin{defi}[Non-locking composition, intuitive definition]

The composition  \(A_p\subst{A_c}{k} = \OA{S}{s_0}{V}{\sigma_0}{J}{T}\) with \(A_c \defobject \OA{S_c}{s_{0c}}{V_c}{\sigma_{0c}}{J_c}{T_c}\), \(A_p \defobject \OA{S_p}{s_{0p}}{V_p}{\sigma_{0p}}{J_p}{T_p}\) and \(k \in J_p\), is a non-locking composition if there is a reachability predicate such that from reachable configurations, either a transition is possible in \(A_p\subst{A_c}{k}\) or there were no possible transition in \(A_p\) to begin with
\begin{multline*}
	\forall s \in S, \forall \sigma: V \to \values, \mpar{\sigma \vdash \reach{A_p\subst{A_c}{k}}\mpar{s}} \implies \mpar{\exists t \in \fOT{s}, \exists \nu: \fvars{t} \to \values, \sigma \uplus \nu \vdash \fguard{t}} \\
	\vee \forall t_p \in \fOT{\pi_{S_p}\mpar{s}}, \forall \nu: \fvars{t_p} \to \values, \sigma \uplus \nu \nvdash \fguard{t_p}
\end{multline*}
\end{defi}

To illustrate the introduction of a deadlock in a composition and the definition of non-locking composition, we consider the following example.
\begin{exi}[Deadlocks introduced by composition]
% TODO: give a context, a circuit cannot block physical time from progressing
\Ludo{TODOludo}{lire l explication quand elle y sera}
\begin{figure}
\input{randomtick_clock.tex}
\vrule
\input{register_anytick.tex}
\caption{On the left: A clock which imposes a tick; On the right: A modified version of the counter at figure \ref{fig:tlh}}
\label{fig:anytick}
\end{figure}
% TODO: explain
\begin{figure}
\input{Composition_deadlock.tex}
\vrule
\input{Composition_nolock.tex}
\caption{On the left: Deadlocks introduced by composition; On the right: No deadlock introduced by composition}
\label{fig:deadlock}
\end{figure}
% TODO: explain
\end{exi}

When expanding the aliases \(S, i, V\) in \Ludo{the definition}{which def? pas lu la suite car je savais pas ou ca allait}, and remarking that the last existential quantifier can be unfolded into a transition in \(A\) and one in \(A_p\) with the added condition that they match, the definition become mildly similar with the deadlock reduction one.
It is not a coincidence because their goal is to caracterise the same kind of compatibility between automata.
Actually it is possible to simplify both definition and make them even more similar.

\begin{defi}[Deadlock reduction, working definition]
A simulation \(\leq\) is deadlock reducing if in all the \(R: S_1 \times S_2 \to \rformulae[V_1 \uplus V_2]\) such that \(\wrel{A_1}{A_2}{R}\), there is one that also satisfies
\begin{multline*}
	\forall \mpar{s_1, s_2} \in S_1 \times S_2, \forall \sigma: V_1 \uplus V_2 \uplus \bigcup_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \to \values, \\
	\sigma \vdash R\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}
\end{multline*}
\Ludo{Question}{on the composed side you might have more variables, eg communicated by the sub-automata or by a sub-hole or stored in the state, how are they quantified?}
\Quentin{Merge-conflict}{I edited the formula above before reading your message, is your question still relevant?}
\end{defi}
\begin{lem}
The intuitive and working definition of deadlock reduction are equivalent.
\end{lem}
